---
title: "Temnos2022Emigration"
author: "Phoebe Koenig"
date: "2023-03-10"
output: html_document
---
### Required Packages

```{r}
library(lme4)
library(nlme)
library(car)
library(ggplot2)
library(plyr)
library(dplyr)
library(MASS)
library(factoextra)
library(performance)
library(emmeans)

```

### **Data**

```{r}


summarydata<-read.csv("~/Temnos2022Emigration/summaryraiddata2022.csv")
```

# Variables

### List:

Here are the variables in the dataset:

**Collected.Date**: Date colony was collected.

**Divided.Date:** Date colony was divided evenly into two matched pairs.

**Pair.Group:** Number identifier for pair group. Each colony collected in the field was divided between 2 nests as a matched pair. Nests with the same Pair group are genetically related (came from the same colony) but one nest in each pair experienced a raiding trial, and one did not.

**Nest:** Number identifier for nest.

**Collection.Site:** Site at which nest containing ant was collected.

**Worker.Count:** Number of workers in collected colony before split. Should be the same for both nests in a pair group.

**Pupae.Count:** Number of pupae in collected colony before split. Should be the same for both nests in a pair group.

**Larvae.Count:** Number of larvae in collected colony before split. Should be the same for both nests in a pair group.

**Queens.Dealate:** Number of dealate queens in collected colony before split. Should be the same for both nests in a pair group.

**Queens.Alate:** Number of alate queens in collected colony before split. Should be the same for both nests in a pair group.

**Males.Alate:** Number of males in collected colony before split. Should be the same for both nests in a pair group.

**Eggs:** Number of eggs in collected colony before split. Should be the same for both nests in a pair group.

**Raided:** Whether or not the nest was raided. Y means yes, N means no.

**Pair.Raid.Date.Start:** Date on which Raiding/nonraided trial began

**Pair.Raid.Date.End:** Date on which Raiding/nonraided trial ended

**Pair.Frozen.Date:** Date on which ants in pair group were frozen

**Raid.Payoff.Pupae:** Net increase in pupae in parasite nest from raiding trial 

**Raid.Payoff.Larvae:** Net increase in larvae in parasite nest from raiding trial 

**Nest.Switch:** Y means ants moved into empty nest. N means they remained in original nest. 

**Notes:** Notes on witnessed raid event and success

**Final.Worker.Count** : The number of workers in the nest on the day it was censused/frozen.

**Final.Queen.Count.Dealate** : The number of queens without wings in the nest on the day it was censused/frozen.

**Final.Pupae.Count** : The number of pupae in the nest on the day it was censused/frozen.

**Final.Larvae.Count** : The number of larvae in the nest on the day it was censused/frozen.

**Final.Egg.Count** : The approximate number of eggs in the nest on the day it was censused/frozen.

**Final.Queen.Count.Alate** : The number of winged queens in the nest on the day it was censused/frozen.

**Final.Male.Count**: The number of males in the nest on the day it was censused/frozen.



```{r}

summarydata$Collected.Date <- as.Date(as.character(summarydata$Collected.Date), format = "%m/ %d/%y")
summarydata$Divided.Date <- as.Date(as.character(summarydata$Divided.Date), format = "%m/ %d/%y")


```



# Nest Switch

```{r}

nestswitch<-subset(summarydata,Nest.Switch!="NA")
nestswitch<-subset(nestswitch,Nest!="31")
nestswitch$Pair.Raid.Date.Start[1]="26-Jul-22"
#Remove Nest 31 because the parasite had no net increase in brood and i did not see them raiding
nestswitch$Pair.Raid.Date.End <- as.Date(as.character(nestswitch$Pair.Raid.Date.End), format = "%d-%b-%y")
nestswitch$Pair.Raid.Date.Start <- as.Date(as.character(nestswitch$Pair.Raid.Date.Start), format = "%d-%b-%y")


#Dataset is successfully loaded but for some random reason missing this date so i'm adding manually
nestswitch$Raid.Duration<-nestswitch$Pair.Raid.Date.End-nestswitch$Pair.Raid.Date.Start
#Had to make a new column that had nest switch in a logical way
nestswitch$Raid.Payoff.Total<-nestswitch$Raid.Payoff.Pupae+nestswitch$Raid.Payoff.Larvae
nestswitch$Final.Pupae.Adult.Ratio<-nestswitch$Final.Pupae.Count/nestswitch$Final.Worker.Count

nestswitch$Proportion.Stolen<-nestswitch$Raid.Payoff.Total/(nestswitch$Final.Larvae.Count+nestswitch$Final.Pupae.Count)
for (p in 1:nrow(nestswitch)) {if (nestswitch$Raided[p]=="N") {nestswitch$Proportion.Stolen[p]=0}}

#Are raided ants significantly more likely to move than unraided ants?
nestmodel<-glmer(Logical.NS~Raided+(1|Pair.Group), data=nestswitch, family="binomial")
summary(nestmodel)
Anova(nestmodel,type="III")


#Is the proportion of ants that moved in the raided group significantly different than what we would expect by chance?
prop.test(12,14,p=0.5)
#The proportion of colonies where the majority moved was .857 CI=.56-.97, which is significantly different than what we would expect by chance (p=0.02, 1-sample test of a proportion)




```

